Hijacking Epidemiology Simulators with Probabilistic Programming

### 2019-04-26 paper/icml_ai4good/main.tex
new abstract

\begin{abstract}

A goal of probabilistic programming is to couple simulators, with inference. This is 
because stochastic simulators are used prominently in many industrial settings,
do not require one to construct hand-crafted joint distributions as they implicitly 
define a joint distribution of the program and encode learned structures 
directly. This makes simulators powerful tools and much of machine learning (ML) and 
Artificial Intelligence (AI)
can be seen as trying to emulate such simulators from a purely data-driven approach.
However, in the 
ML/AI setting, although we can often infer outcomes, we have little understanding about what 
in the data led to the outputted inferences. 
This makes it challenging to deploy ML/AI systems into the wild, especially in health-related and safety-critical domains, such
as epidemiology, as we lose \emph{interpretability}. 
In this work, we explain how to design ML/AI systems that combine
probabilistic programming systems (PPSs) and epidemiology simulators, to extract
fully interpretable structures, enabling policy makers 
and practitioners to make decisions from simulator outputs that can be fully reasoned on.
 % interpretable inferences. 
In particular, we demonstrate this for the Malaria disease in two commonly used simulators; EMOD and OpenMalaria.
\end{abstract}

@@

Critically, this hijacking can be performed on a wide variety of simulators
with only minimal changes to the simulator.  Our framework already
supports 13 popular languages for the base simulator, and can easily
be extended to others if needed.  Moreover, though our focus
is on hijacking epidemiology simulators, our general approach and
framework can be applied more widely to a general class of stochastic
simulators.  In time, we hope our approach will play a critical role in
bringing recent advancements in probabilistic programming to bear on
the vast array of existing simulators used throughout the sciences,
thereby providing wide-ranging impacts across a number of fields.

>>>>

\begin{abstract}
Epidemiology simulations have become a fundamental tool in the fight
against the epidemics of various infectious diseases like AIDS and
malaria.
However, the complicated and stochastic nature of these simulators
can mean they are difficult to work with and extract the key information
required by policy makers.
In this paper, we introduce an approach to \emph{hijack} these
simulators with an external
probabilistic programming system (PPS) that controls
their internal random number generator.
This allows for various advanced statistical analyses to be performed
on the simulators in an automated fashion, while also tracking
the behavior of the simulator to uncover key information such
as stochastic structure and generate result summaries.
Critically, this hijacking can be performed with only minimal changes
to the original simulator, allowing for straightforward application
of our approach to a wide variety of contexts and existing simulators.
%
%A goal of probabilistic programming is to couple simulators, with inference. This is 
%because stochastic simulators are used prominently in many industrial settings,
%do not require one to construct hand-crafted joint distributions as they implicitly 
%define a joint distribution of the program and encode learned structures 
%directly. This makes simulators powerful tools and much of machine learning (ML) and 
%Artificial Intelligence (AI)
%can be seen as trying to emulate such simulators from a purely data-driven approach.
%However, in the 
%ML/AI setting, although we can often infer outcomes, we have little understanding about what 
%in the data led to the outputted inferences. 
%This makes it challenging to deploy ML/AI systems into the wild, especially in health-related and safety-critical domains, such
%as epidemiology, as we lose \emph{interpretability}. 
%In this work, we explain how to design ML/AI systems that combine
%probabilistic programming systems (PPSs) and epidemiology simulators, to extract
%fully interpretable structures, enabling policy makers 
%and practitioners to make decisions from simulator outputs that can be fully reasoned on.
% % interpretable inferences. 
%In particular, we demonstrate this for the Malaria disease in two commonly used simulators; EMOD and OpenMalaria.
\end{abstract}

@@

Critically, this hijacking can be performed on a wide variety of simulators
with only minimal changes.  Our framework already
supports 13 popular languages for the base simulator, and can easily
be extended to others if needed.  Moreover, though our focus
is on hijacking epidemiology simulators, our general approach and
framework can be applied more widely to a general class of stochastic
simulators.  In time, we hope our approach will play a critical role in
bringing recent advancements in probabilistic programming to bear on
the vast array of existing simulators used throughout the sciences,
thereby providing wide-ranging impacts across a number of fields.
